---
title: "Working with sf objects"
author: "Emma-Marie Vitskov Jørgensen"
date: "04-February-2022 updated `r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
---

```{r setup, include=FALSE}
library(knitr)
library(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(echo=TRUE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

#  Goals: 

1. In the first exercise, you will learn to construct maps with `tmap()` by modifying the provided code. The libraries, data, and initial code is provided below.
2. In the second exercise, you learn to: 

 - Understand the provided datasets
 - Learn how to reproject spatial data
 - Limit your data into an area of interest
 - Create a new map
3. In the third exercise, you reproject layers to a shared SRS.
4. In the fourth (and optional) exercise, you create a map of Denmark using online data.

#### Required R libraries

We will use the `sf`, `raster`, and `tmap` packages.
Additionally, we will use the `spData` and `spDataLarge` packages that provide new datasets. To install the `spDataLarge`, please load it directly from the author's github repository at https://github.com/Nowosad/spDataLarge. If the line in the chunk below does not work, check the repository for the other options.

```{r}
library(sf)
library(raster)
library(tmap)
library(spData)

# To install the spDataLarge uncomment the next line

#install.packages("spDataLarge", repos = "https://geocompr.r-universe.dev")
#install.packages("spData")
install.packages("devtools")
devtools::install_github("geocompr/geocompkg")
library(spDataLarge)
```


#  Exercise 1 - Make a map

In this exercise you will learn to make a map with `tmap` library, by adding spatial data layers and modifying the arguments that specify their rendering

#### Data sets  

We will use two data sets: `nz_elev` and `nz`. They are contained by the libraries
The first one is an elevation raster object for the New Zealand area, and the second one is an sf object with polygons representing the 16 regions of New Zealand.

#### Existing code

Here is the code to create a new map of New Zealand (following Robin Lovelace's example). Your role is to improve this map based on the suggestions below.

```{r initial-map}
tm_shape(nz_elev)  +
  tm_raster(title = "elev", 
            style = "cont",
            palette = "BuGn") +
  tm_shape(nz) +
  tm_borders(col = "red", 
             lwd = 3) +
  tm_scale_bar(breaks = c(0, 100, 200),
               text.size = 1) +
  tm_compass(position = c("LEFT", "center"),
             type = "rose", 
             size = 2) +
  tm_credits(text = "A. Sobotkova, 2023") +
  tm_layout(main.title = "My map",
            bg.color = "orange",
            inner.margins = c(0, 0, 0, 0))
```


## Tasks 

1. Change the map title from "My map" to "New Zealand".
2. Update the map credits with your own name and today's date.
3. Change the color palette to "-RdYlGn". 
  (You can also try other palettes from http://colorbrewer2.org/)
4. Put the north arrow in the top right corner of the map.
5. Improve the legend title by adding the used units (m asl).
6. Increase the number of breaks in the scale bar.
7. Change the borders' color of the New Zealand's regions to black. 
  Decrease the line width.
8. Change the background color to any color of your choice.


## Your solution

```{r improved-map}
tm_shape(nz_elev)  +
  tm_raster(title = "m asl", 
            style = "cont",
            palette = "-RdYlGn") +
  tm_shape(nz) +
  tm_borders(col = "black", 
             lwd = 1) +
  tm_scale_bar(breaks = c(0, 100, 200, 300),
               text.size = 1) +
  tm_compass(position = c("RIGHT", "top"),
             type = "rose", 
             size = 2) +
  tm_credits(text = "E. V. Jørgensen, 21.02.2023") +
  tm_layout(main.title = "New Zealand",
            bg.color = "beige",
            inner.margins = c(0, 0, 0, 0))
```


# Exercise 2 - Inspect and display spatial data

We will use two data sets: `srtm` and `zion` , explore them and make a map.
The first one is an elevation raster object for the Zion National Park area in Israel, and the second one is an sf object with polygons representing borders of the Zion National Park.

Read two new datasets, `srtm` and `zion`, using the code below and then follow the instructions and answer the questions related.

####  Preparation: Load data from provided package
You need to have installed the `spDataLarge` package for the following code to work. 

```{r load-data}
srtm  <-  raster(system.file("raster/srtm.tif", package = "spDataLarge"))
zion  <-  read_sf(system.file("vector/zion.gpkg", package = "spDataLarge"))
```

## Instructions and questions
Type answers to the questions as code comments next to or under the code used

* Display the `zion` object and view its structure.

  - What can you say about the content of this file?
  - What type of data does it store? 
  - What is the coordinate system used?
  - How many attributes does it contain?
  - What is its geometry?

* Display the `srtm` object and view its structure.

  - What can you say about the content of this file? 
  - What type of data does it store?
  - What is the coordinate system used? 
  - How many attributes does it contain?
  - How many dimensions does it have? 
  - What is the data resolution?
 
## Your solution and answers

Display the `zion` object

```{r inspect-data-1}
zion
# The CRS is UTM Zone 12, Northern Hemisphere. It has 1 feature. 
# The values of the coordinate system are xmin: 302903.1 ymin: 4112244 xmax: 334735.5 ymax: 4153087. 

class(zion)
# The zion data is a data frame and a tibble. It is also a SF object.

st_crs(zion)
# The coordinate reference system used is GRS80.

library(tidyverse)
view(zion)
zion$geom
#I think zion is a projection (using northing and southing), because the values of the x- and y-axis' are so big. 
 

zion
# I don't know how many attributes does it contain??? 

zion
# the geometry of Zion is a polygon.

```

Displaying the `srtm` object

```{r inspect-data-2}
srtm
# srtm is a Rasterlayer. Its CRS is  +proj=longlat +datum=WGS84 +no_defs. The dimensions of the raster is 457 rows and 465 columns. 

class(srtm)
# it stores raster data. 

crs(srtm)
# The coordinate system used is a 3D geographic coordinate system (long/lat). 

srtm
# How many attributes does it contain???

srtm
# The data has two dimensions: 457 rows and 465 columns

srtm
# The data resolution is 0.0008333333, 0.0008333333  (x, y)
```
I failed to find the attributes of both `zion` and `srtm`. How are they found, what do they mean and what are they used for?


# Exercise 3 - Reconciling Coordinate systems

Now that you know that coordinate systems differ, make them compatible twice over, converting to one and then another CRS!


## Instructions

* Reproject the `srtm` dataset into the coordinate reference system used in the `zion` object. 

  - Create a new object `srtm2` with the `zion` crs. You can label it `srtm_####` writing the EPSG out for easy differentiation.
  - Visualize the results using the `plot()` function.

* Reproject the `zion` dataset into the coordinate reference system used in the `srtm` object.
  - Create a new object `zion2` or `zion_####` with the `srtm` crs.
  - Visualize the results (`zion2` together with `srtm`) using the `plot()` function.

## Your solution

Reprojecting the `srtm` dataset

```{r reproject-srtm}
library(sf)
library(raster)

#Get the crs of zion and srtm
st_crs(zion, asText=TRUE)#the crs is UTM Zone 12, Northern Hemisphere
crs(srtm) #the crs is +proj=longlat +datum=WGS84 +no_defs

#create a new object containing the same information as srtm (for now)
srtm2 <- srtm


#create a new variable containing the crs of zion

#the_crs <- CRS('+init=EPSG:16012')
#the_crs <- "UTM Zone 12, Northern Hemisphere"
#the_crs <- st_crs(zion, asText = TRUE) 

# Project srtm to match the CRS of zion
#srtm2_crs <- projectRaster(srtm2, crs = the_crs, method = "ngb")
srtm2_crs <- projectRaster(srtm2, crs = 4326, method = "ngb")

crs(srtm2) <- srtm2_crs


crs(srtm2)


### Vizualisation ###
plot(st_geometry(zion2))
plot(srtm, add = TRUE)


```

* Use `st_transform()` to transform the vector `zion` object to the CRS in `the_crs`.
* Use `projectRaster()` to transform the raster `srtm` object to the CRS in `the_crs`. This will take a few seconds.

Now I will reproject the `zion` dataset. 

* Reproject the `zion` dataset into the coordinate reference system used in the `srtm` object.
  - Create a new object `zion2` or `zion_####` with the `srtm` crs.
  - Visualize the results (`zion2` together with `srtm`) using the `plot()` function.

```{r reproject-zion}
#find crs
st_crs(zion)#the crs is UTM Zone 12, Northern Hemisphere
crs(srtm) #the crs is +proj=longlat +datum=WGS84 +no_defs

# assigning the crs of srtm to an object
crs_2 <-"+proj=longlat +datum=WGS84 +no_defs"

# creating a zion2 object
zion2 <- zion

# setting the crs of zion2 to be the crs_2
st_crs(zion2) <- crs_2

#checking the crs of zion2
st_crs(zion2)
# the crs is now +proj=longlat +datum=WGS84 +no_defs


# Visualization
plot(st_geometry(zion2))
plot(srtm, add = TRUE)

```
 
 Now the crs's of zion and srtm match!

## Refs
Tennekes, Martijn. 2019. Tmap: Thematic Maps. https://CRAN.R-project.org/package=tmap.

